Visualizing probability distributions across bivariate cyclic temporal granularities

### 2020-03-19 main.Rmd
till cyclic calendar algebra

<!-- some discussions on primary units of calendar -->
**Example:** So far we have used example of cyclic granularities from Gregorian calendar as it is the most widely used calendar. But it is far from being the only one. <!--All calendars fall under three types - solar, lunar or lunisolar/solilunar-->The day is the basic unit of time underlying most calendars [@Reingold2001-kf]. Various calendars, however, use different conventions to structure days into larger units: weeks, months, years and cycle of years. The French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds". Nevertheless, for any calendar a hierarchy table can be defined. For example, in Mayan calendar, one day was referred to as 1 kin and the calendar was structured such that 1 kin = 1 day; 1 uinal = 20 kin; 1 tun  = 18 uinal; 1 katun = 20 tun and 1 baktun = 20 katun

Thus, the hierarchy table for the Mayan calendar would look like the following:

|  G	        |C                 |P
|------------	|------------------|------------|
|  kin      	|kin-of-uinal      |  20        |
|  uinal    	|uinal-of-tun      |  18  	    |
|  tun      	|tun-of-katun      |  20	      |
|  katun    	|katun-of-baktun   |  20        |
|  baktun   	| 1                |	1         |

@@

Circular single-order-up granularities can be used recursively to obtain multiple order up circular granularity. Since, the operation requires the use of modular arithmetic, it is important that the label mapping of the individual circular single order-up granularity is an identity function, that is, $L(x) = x\quad \forall x$. The label mapping of the resultant multiple-order-up granularity can however be chosen arbitrarily, depending on the context.

@@

@@

\begin{equation} \label{eq8}
\begin{split}
C_{uinal, baktun}(z)  = & L(C_{uinal, tun}(z) + P(uinal, tun)C_{tun,katun}(z) + C_{uinal, katun}C_{katun, baktun}(z)) \\
               = & L(\lfloor z/20\rfloor \mod 18 + 20*\lfloor z/20*18\rfloor \mod 20  \\
               &~~~ + 20*18*20\lfloor z/20*18*20\rfloor \mod 20) \\
\end{split}
\end{equation}

@@

We consider the case of only one quasi-circular single order-up granularity in the hierarchy table. Any multiple order-up quasi-circular granularity $C_{l, m}(z)$ can then be considered to be a function of two single order-up granularities $C_{l,m'}(z)$ and $C_{m',m}(z)$. Two different approaches need to be employed for the following cases:

@@

**Example:** Let us revisit Gregorian calendar for addressing this case. Suppose we have a hierarchy using linear granularties from Gregorian calendar. Since months consists of unequal number of days, any linear granularity with higher order than months will also have unequal number of days. This is an example of a hierarchy structure which has circular and quasi-circular single-order-up granularities. Only one single-order-up granularity day-of-month is quasi-circular.

|  G	        |C                 |P
|------------	|------------------|----------------|
|  minute     |minute-of-hour    |  60            |
|  hour    	  |hour-of-day       |  24  	        |
|  day      	|day-of-month      |quasi-circular  |
|  month    	|month-of-year     |  12            |
|  year   	  | 1                |	1             |

@@

For a hierarchy table $H_n: (G, C, k)$ with $l_1, l_2, m_1, m_2 \in {1, 2, \dots, n}$ and $l_2<l_1$ and $m_2>m_1$, we have

\begin{equation} \label{eq:all-circular-multiple}
C_{G_{l1}, G_{m1}}(z) = C_{G_{l2}, G_{m2}}(\lfloor z/k(l_2,l_1) \rfloor\mod k(m_1, m_2))
\end{equation}

**Example:** Considering the same example of Mayan Calendar, it is possible to compute the single-order-up granularity tun-of-katun  given the multiple-order-up granularity uinal-baktun using equation \ref{eq:all-circular-multiple}

>>>>

<!-- some discussions on primary units of calendar -->
**Example:** So far we have used example of cyclic granularities from Gregorian calendar as it is the most widely used calendar. But it is far from being the only one. <!--All calendars fall under three types - solar, lunar or lunisolar/solilunar-->The day is the basic unit of time underlying most calendars [@Reingold2001-kf]. Various calendars, however, use different conventions to structure days into larger units: weeks, months, years and cycle of years. The French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds". Nevertheless, for any calendar a hierarchy table can be defined. For example, in Mayan calendar, one day was referred to as 1 kin and the calendar was structured such that 1 kin = 1 day; 1 uinal = 20 kin; 1 tun  = 18 uinal; 1 katun = 20 tun and 1 baktun = 20 katun. \autoref{tab:tab-mayan} represents the the hierarchy table for the Mayan calendar.

```{r tab-mayan}
table_mayan <- tibble(G = c("kin", "uinal", "tun", "katun", "baktun"),
                      C = c("kin-of-uinal","uinal-of-tun"," tun-of-katun", "katun-of-baktun", 1),
                      K = c(20, 18, 20, 20, 1)) 
kable(table_mayan, 
      caption = "Hierarchy table for Mayan calendar. $G$ and $C$ represent the linear and single order-up cyclic granularities respectively. All cyclic granularities are circular with $K$ as the period length.")
```

@@

Circular single-order-up granularities can be used recursively to obtain multiple order up circular granularity. Since, the operation requires the use of modular arithmetic, it is important that the label mapping of the individual circular single order-up granularity is an identity function, that is, $L(x) = x\quad \forall x$. The label mapping of the resultant multiple-order-up granularity can however be chosen depending on the context.

@@

, for $l < m + 1$ and $P(i, j) = 1, \quad \forall i = j \in \{0, 1, \dots, m-l-1\}$

@@

\begin{equation} \label{eq8}
\begin{split}
C_{uinal, baktun}(z)  = & L(C_{uinal, tun}(z) + P(uinal, tun)C_{tun,katun}(z) + P(uinal, katun)C_{katun, baktun}(z)) \\
               = & L(\lfloor z/20\rfloor \mod 18 + 18*\lfloor z/20*18\rfloor \mod 20  \\
               &~~~ + 18*20\lfloor z/20*18*20\rfloor \mod 20) \\
\end{split}
\end{equation}

@@

We consider the case of only one quasi-circular single order-up granularity in the hierarchy table. Any multiple order-up quasi-circular granularity $C_{l, m}(z)$ can then be obtained as a function of a single (or multiple) order-up circular granularity $C_{l,m'}(z)$ and single order-up quasi-circular granularity $C_{m',m}(z)$ or vice versa. Two different approaches need to be employed for the following cases:

@@

**Example:** <!--Let us revisit Gregorian calendar for addressing this case.--> Suppose we have a hierarchy using linear granularties from Gregorian calendar. Since months consists of unequal number of days, any linear granularity with higher order than months will also have unequal number of days. This is an example of a hierarchy structure which has both circular and quasi-circular single-order-up granularities. Only one single-order-up granularity day-of-month is quasi-circular.

|  G	        |C                 |K
|------------	|------------------|----------------|
|  minute     |minute-of-hour    |  60            |
|  hour    	  |hour-of-day       |  24  	        |
|  day      	|day-of-month      | k(day, month)  |
|  month    	|month-of-year     |  12            |
|  year   	  | 1                |	1             |

@@

For a hierarchy table $H_n: (G, C, K)$ with $l_1, l_2, m_1, m_2 \in {1, 2, \dots, n}$ and $l_2<l_1$ and $m_2>m_1$, we have

\begin{equation} \label{eq:all-circular-multiple}
C_{G_{l_1}, G_{m_1}}(z) = C_{G_{l_2}, G_{m_2}}(\lfloor z/k(l_2,l_1) \rfloor\mod P(m_1, m_2))
\end{equation}

**Example:** Considering the same example of Mayan Calendar, it is possible to compute the single-order-up granularity tun-of-katun  given the multiple-order-up granularity uinal-baktun using equation \ref{eq:all-circular-multiple}.
